---
title: "pPPCA_model_validation_TBPACTS"
output: html_document
---
## Get pkgs
```{r}
pkgs<-c("tidyverse","lme4","ggplot2","emmeans","RColorBrewer","patchwork","gt",
          "grid","broom.mixed","pbkrtest","jtools","car")
lapply(pkgs, require, character.only=T)
```

## Read in data
```{r}
## write out of dose-response models in R
df<-read.csv("/Users/mairi.mcclean/Desktop/btz/manuscript_posters/CPT_PSP_submission/submitted/revisions/extra_analysis/processed_data/260526mm_pPCA_TB1014.csv")
df<-df|>
  mutate(across(c(USUBJID, ARM,SEX), as.factor))
```

### Quick vis
```{r}
bm_all<-df |>
  pivot_longer(c(logCFU,logTTP,pc1),
    names_to="assay",
    values_to="value"
  ) |>
  select(-c(SEX,AGE,pc2))
```


```{r}
ggplot(data = bm_all, aes(x = DAY, y = value, color = USUBJID)) +
  geom_line() +
  stat_summary(
    fun = median,
    geom = "line",
    aes(group = assay, color=assay),
    size = 1.5
  ) +
  scale_color_manual(
    values = c(
      setNames(rep("grey70", length(unique(bm_all$USUBJID))), unique(bm_all$USUBJID)),
      setNames(brewer.pal(4, "Set2"), unique(bm_all$assay))
    )
  ) +
  facet_wrap(
    ~ assay,
    scales = "free_y",
    ncol = 4,
             labeller = labeller(
              assay = c("logCFU" = "log10 CFU/ml",
                        "logTTP" = "log10 MGIT TTP (h)",
                        "pc1_all"="PC1")
            )) +
  ggtitle("Per-patient trajectories by measurement") +
  theme_minimal() +
  theme(panel.grid.major = element_blank(),
        legend.position = "none",
        axis.title.x = element_text(size=12),
        axis.title.y = element_text(size=12),
        title = element_text(size = 15),
        legend.text = element_text(size = 15),
        strip.text=element_text(size=12)) +
  labs(y = "Assay value", x = "Days on treament")
```


## Models
### Linear mixed models
#### TTP
```{r}
### Random slope + INTERCEPT
LinearMixed_TTP <- lmer(logTTP ~ DAY:(1+ARM) + (DAY|USUBJID),
                      data = df) 
coef_model_TTP <- coef(LinearMixed_TTP)$USUBJID

# Now tidy using broom
stage2_slope_ttp <- tidy(LinearMixed_TTP) %>%
  filter(effect == "fixed") %>%
  mutate(ci95 = str_c(
    format(round(estimate,digits=4),nsmall=2),
    " (",
    format(round(estimate - 1.96*std.error,digits=4),nsmall=2),
    ", ",
    format(round(estimate + 1.96*std.error,digits=4),nsmall=2),
    ")"
    )
    
  ) %>%
  dplyr::select(term,ci95)

stage2_slope_ttp$se_fixed_effects <- summary(LinearMixed_TTP)$coefficients[, "Std. Error"]
confint(LinearMixed_TTP, method = "Wald")
# For comparison 
comp_ttp_coefs<- stage2_slope_ttp
```

```{r}
summary(LinearMixed_TTP)
```

```{r}
gr_ttp <- ref_grid(LinearMixed_TTP, cov.keep= c('ARM', 'DAY'))
emm_ttp <- emmeans(gr_ttp, spec= c('ARM', 'DAY'), level= 0.95)
emm_ttp_df <- as.data.frame(emm_ttp)
```

```{r}
ggplot(df, aes(x=DAY, y=logTTP))+
        geom_line(data= data.frame(emm_ttp), aes(y= emmean), color ='blue', size = 1.5) +
        scale_x_continuous(limits = c(0, 15), breaks = c(0,2,4,6,8,10,12,14))+
        facet_wrap(~ARM,nrow=2)+
        labs(y = "log10(TTP) (hrs)",
             x = "Day(s) of treatment")+
        theme(axis.title = element_text(size = 16))

```

#### CFU
```{r}
LinearMixed_CFU <- lmer(logCFU ~ DAY:(1+ARM) + (DAY|USUBJID),
                      data = df) 
coef_model_CFU <- coef(LinearMixed_CFU)$USUBJID

# Now tidy using broom
stage2_slope_cfu <- tidy(LinearMixed_CFU) %>%
  filter(effect == "fixed") %>%
  mutate(ci95 = str_c(
    format(round(estimate,digits=4),nsmall=2),
    " (",
    format(round(estimate - 1.96*std.error,digits=4),nsmall=2),
    ", ",
    format(round(estimate + 1.96*std.error,digits=4),nsmall=2),
    ")"
    )
    
  ) %>%
  dplyr::select(term,ci95)

stage2_slope_cfu$se_fixed_effects <- summary(LinearMixed_CFU)$coefficients[, "Std. Error"]
confint(LinearMixed_CFU, method = "Wald")
# For comparison 
comp_cfu_coefs<- stage2_slope_cfu
```

```{r}
summary(LinearMixed_CFU)
```

```{r}
gr_cfu <- ref_grid(LinearMixed_CFU, cov.keep= c('ARM', 'DAY'))
emm_cfu <- emmeans(gr_cfu, spec= c('ARM', 'DAY'), level= 0.95)
emm_cfu_df <- as.data.frame(emm_cfu)
```

```{r}
ggplot(df, aes(x=DAY, y=logCFU))+
        geom_line(data= data.frame(emm_cfu), aes(y= emmean), color ='blue', size = 1.5) +
        scale_x_continuous(limits = c(0, 15), breaks = c(0,2,4,6,8,10,12,14))+
        facet_wrap(~ARM,nrow=2)+
        labs(y = "log10(CFU)",
             x = "Day(s) of treatment")+
        theme(axis.title = element_text(size = 16))

```

#### PC1
```{r}
LinearMixed_PC1 <- lmer(pc1 ~ DAY:(1+ARM) + (DAY|USUBJID),
                      data = df) 
coef_model_pc1 <- coef(LinearMixed_PC1)$USUBJID

# Now tidy using broom
stage2_slope_pc1 <- tidy(LinearMixed_PC1) %>%
  filter(effect == "fixed") %>%
  mutate(ci95 = str_c(
    format(round(estimate,digits=4),nsmall=2),
    " (",
    format(round(estimate - 1.96*std.error,digits=4),nsmall=2),
    ", ",
    format(round(estimate + 1.96*std.error,digits=4),nsmall=2),
    ")"
    )
    
  ) %>%
  dplyr::select(term,ci95)

stage2_slope_pc1$se_fixed_effects <- summary(LinearMixed_PC1)$coefficients[, "Std. Error"]
confint(LinearMixed_PC1, method = "Wald")
# For comparison 
comp_pc1_coefs<- stage2_slope_pc1
```

```{r}
summary(LinearMixed_PC1)
```

```{r}
gr_pc <- ref_grid(LinearMixed_PC1, cov.keep= c('ARM', 'DAY'))
emm_pc <- emmeans(gr_pc, spec= c('ARM', 'DAY'), level= 0.95)
emm_pc_df <- as.data.frame(emm_pc)
```

```{r}
ggplot(df, aes(x=DAY, y=pc1))+
        geom_line(data= data.frame(emm_ttp), aes(y= emmean), color ='blue', size = 1.5) +
        scale_x_continuous(limits = c(0, 15), breaks = c(0,2,4,6,8,10,12,14))+
        facet_wrap(~ARM,nrow=2)+
        labs(y = "log10(TTP) (hrs)",
             x = "Day(s) of treatment")+
        theme(axis.title = element_text(size = 16))

```

```{r}
summary(LinearMixed_CFU)
colnames(coef_model_pc1)
```

```{r}
colnames(coef_model_pc1)
```

#### Forest plot
```{r}
plt <- plot_coefs(LinearMixed_PC1,LinearMixed_TTP,LinearMixed_CFU,
           model.names = c('PC1 [CFU / MGIT TTP] (n=105)',
                           'MGIT TTP [liquid culture] (n=105)',
                           'CFU [solid culture] (n=105)'),
          omit.coefs = "(Intercept)",
           coefs=c('CLOFAZIMINE'='DAY:ARMCLOFAZIMINE',
                   'PYRAZINAMIDE'='DAY:ARMPYRAZINAMIDE',
                   'Rifafour e-275'='DAY:ARMRifafour e-275',
                   'TMC207 + PA-824 + CLOFAZIMINE'='DAY:ARMTMC207 + PA-824 + CLOFAZIMINE',
                   'TMC207 + PA-824 + PYRAZINAMIDE'='DAY:ARMTMC207 + PA-824 + PYRAZINAMIDE',
                   'TMC207 + PA-824 + PYRAZINAMIDE + CLOFAZIMINE'='DAY:ARMTMC207 + PA-824 + PYRAZINAMIDE + CLOFAZIMINE',
                   'TMC207 + PYRAZINAMIDE + CLOFAZIMINE'='DAY:ARMTMC207 + PYRAZINAMIDE + CLOFAZIMINE'
                   ),
           scale=TRUE, size=15,legend.title="")

plt <- plt +
  scale_color_manual(values = c(
    "CFU [solid culture] (n=105)" = "#F8766D",
    "MGIT TTP [liquid culture] (n=105)" = "#7CAE00",
    "PC1 [CFU / MGIT TTP] (n=105)" = "#00BFC4"
  )) +
  facet_wrap(~ model, ncol=2, scales = "free_x") +
  theme_minimal() +
#+   scale_x_continuous(breaks = c(0, seq(-1, 1, by = 0.1))) + 
  theme(axis.text.x = element_text(size = 9),
        axis.title.x=element_text(size=12),
        strip.text=element_text(size=10),
        axis.text.y = element_text(size = 10),
#        axis.title.y=element_blank(),
        legend.text=element_text(),
        legend.title = element_text(),
        legend.position='none'
        )+
  guides(shape=guide_legend(reverse = TRUE))+
  labs(x='Model coefficient',
       y='Treatment arm')
#CFU-'#F8766D'
#TTP-'#7CAE00'
#PC1-'#00BFC4'
plt

#ggsave("260115_model_coefs_OG_models_PC1.tiff", plt, device='tiff', 
#       width =178,dpi = 300, bg='white',units='mm' ) 
```


#### ANOVA of model estimates
##### TTP
```{r}
anova(LinearMixed_TTP, type=3)
```

##### CFU
```{r}
anova(LinearMixed_CFU, type=3)
```

##### PC1
```{r}
anova(LinearMixed_PC1, type=3)
```



#### ANOVA of patient estimates
```{r}
## get patient level slopes per measurement for ANOVA
write.csv(coef_model_pc1, "PC_model_est_per_px.csv")
write.csv(coef_model_CFU, "CFU_model_est_per_px.csv")
```

##### TTP
```{r}
write.csv(coef_model_TTP, "TTP_model_est_per_px.csv")
idv_estimates <- read_csv("TTP_model_est_per_px.csv", col_names = TRUE)
names(idv_estimates)[1] <- "Participant.ID"
idv_estimates$USUBJID <- as.factor(idv_estimates$Participant.ID)

```
                                
```{r}
tx_alloc<-df|>
  select(c(USUBJID,ARM))
patient_TTP_estimates <- left_join(idv_estimates,tx_alloc, by = 'USUBJID') %>%
  mutate(ch1 = ifelse(ARM == "CLOFAZIMINE", 1, 0),
         ch2 = ifelse(ARM == "PYRAZINAMIDE", 1, 0),
         ch3 = ifelse(ARM == "Rifafour e-275", 1, 0),
         ch4 = ifelse(ARM == "TMC207 + PA-824 + CLOFAZIMINE", 1, 0),
         ch5 = ifelse(ARM == "TMC207 + PA-824 + PYRAZINAMIDE", 1, 0),
         ch6 = ifelse(ARM == "TMC207 + PA-824 + PYRAZINAMIDE + CLOFAZIMINE", 1, 0),
         ch7 = ifelse(ARM == "TMC207 + PYRAZINAMIDE + CLOFAZIMINE", 1, 0),
         ) %>%
  mutate(idv_slp = DAY + 
           `DAY:ARMCLOFAZIMINE`*ch1 + 
           `DAY:ARMPYRAZINAMIDE`*ch2+
           `DAY:ARMRifafour e-275`*ch3+
           `DAY:ARMTMC207 + PA-824 + CLOFAZIMINE`*ch4+
           `DAY:ARMTMC207 + PA-824 + PYRAZINAMIDE`*ch5+
           `DAY:ARMTMC207 + PA-824 + PYRAZINAMIDE + CLOFAZIMINE`*ch6+
           `DAY:ARMTMC207 + PYRAZINAMIDE + CLOFAZIMINE`*ch7)

write.csv(patient_TTP_estimates,
          "/Users/mairi.mcclean/Desktop/btz/manuscript_posters/CPT_PSP_submission/submitted/revisions/extra_analysis/processed_data/TTP_slopes_per_px.csv")
```


##### CFU
```{r}
write.csv(coef_model_CFU, "CFU_model_est_per_px.csv")
idv_estimates <- read_csv("CFU_model_est_per_px.csv", col_names = TRUE)
names(idv_estimates)[1] <- "Participant.ID"
idv_estimates$USUBJID <- as.factor(idv_estimates$Participant.ID)
```
                                
```{r}
tx_alloc<-df|>
  select(c(USUBJID,ARM))
patient_CFU_estimates <- left_join(idv_estimates,tx_alloc, by = 'USUBJID') %>%
  mutate(ch1 = ifelse(ARM == "CLOFAZIMINE", 1, 0),
         ch2 = ifelse(ARM == "PYRAZINAMIDE", 1, 0),
         ch3 = ifelse(ARM == "Rifafour e-275", 1, 0),
         ch4 = ifelse(ARM == "TMC207 + PA-824 + CLOFAZIMINE", 1, 0),
         ch5 = ifelse(ARM == "TMC207 + PA-824 + PYRAZINAMIDE", 1, 0),
         ch6 = ifelse(ARM == "TMC207 + PA-824 + PYRAZINAMIDE + CLOFAZIMINE", 1, 0),
         ch7 = ifelse(ARM == "TMC207 + PYRAZINAMIDE + CLOFAZIMINE", 1, 0),
         ) %>%
  mutate(idv_slp = DAY + 
           `DAY:ARMCLOFAZIMINE`*ch1 + 
           `DAY:ARMPYRAZINAMIDE`*ch2+
           `DAY:ARMRifafour e-275`*ch3+
           `DAY:ARMTMC207 + PA-824 + CLOFAZIMINE`*ch4+
           `DAY:ARMTMC207 + PA-824 + PYRAZINAMIDE`*ch5+
           `DAY:ARMTMC207 + PA-824 + PYRAZINAMIDE + CLOFAZIMINE`*ch6+
           `DAY:ARMTMC207 + PYRAZINAMIDE + CLOFAZIMINE`*ch7)

write.csv(patient_CFU_estimates,
          "/Users/mairi.mcclean/Desktop/btz/manuscript_posters/CPT_PSP_submission/submitted/revisions/extra_analysis/processed_data/CFU_slopes_per_px.csv")
```


##### PC1
```{r}
write.csv(coef_model_pc1, "PC_model_est_per_px.csv")
idv_estimates <- read_csv("PC_model_est_per_px.csv", col_names = TRUE)
names(idv_estimates)[1] <- "Participant.ID"
idv_estimates$USUBJID <- as.factor(idv_estimates$Participant.ID)

```
                                
```{r}
tx_alloc<-df|>
  select(c(USUBJID,ARM))
patient_PC1_estimates <- left_join(idv_estimates,tx_alloc, by = 'USUBJID') %>%
  mutate(ch1 = ifelse(ARM == "CLOFAZIMINE", 1, 0),
         ch2 = ifelse(ARM == "PYRAZINAMIDE", 1, 0),
         ch3 = ifelse(ARM == "Rifafour e-275", 1, 0),
         ch4 = ifelse(ARM == "TMC207 + PA-824 + CLOFAZIMINE", 1, 0),
         ch5 = ifelse(ARM == "TMC207 + PA-824 + PYRAZINAMIDE", 1, 0),
         ch6 = ifelse(ARM == "TMC207 + PA-824 + PYRAZINAMIDE + CLOFAZIMINE", 1, 0),
         ch7 = ifelse(ARM == "TMC207 + PYRAZINAMIDE + CLOFAZIMINE", 1, 0),
         ) %>%
  mutate(idv_slp = DAY + 
           `DAY:ARMCLOFAZIMINE`*ch1 + 
           `DAY:ARMPYRAZINAMIDE`*ch2+
           `DAY:ARMRifafour e-275`*ch3+
           `DAY:ARMTMC207 + PA-824 + CLOFAZIMINE`*ch4+
           `DAY:ARMTMC207 + PA-824 + PYRAZINAMIDE`*ch5+
           `DAY:ARMTMC207 + PA-824 + PYRAZINAMIDE + CLOFAZIMINE`*ch6+
           `DAY:ARMTMC207 + PYRAZINAMIDE + CLOFAZIMINE`*ch7)

write.csv(patient_PC1_estimates,
          "/Users/mairi.mcclean/Desktop/btz/manuscript_posters/CPT_PSP_submission/submitted/revisions/extra_analysis/processed_data/PC1_slopes_per_px.csv")
```


#### Simple stats on model slopes
```{r}
#### Original authors ran ANOVA (on patient slopes? Or model estimates) TBD
res.ttp<-aov(idv_slp ~ ARM, data=patient_TTP_estimates)
summary(res.ttp)
```

```{r}
res.cfu<-aov(idv_slp ~ ARM, data=patient_CFU_estimates)
summary(res.cfu)
```

```{r}
res.pc1<-aov(idv_slp ~ ARM, data=patient_PC1_estimates)
summary(res.pc1)
```
> PC1 better than CFU, not better than TTP, to determine an overall effect

##### H0 tests
> How are the H0 tests for beta==0, and control == experimental arm
>> Get all 10 and then correct for MTC?

```{r}
# 'DAY:ARMCLOFAZIMINE=0'
linearHypothesis(LinearMixed_TTP, c('DAY:ARMTMC207 + PA-824 + PYRAZINAMIDE=DAY:ARMRifafour e-275'))
```

```{r}
linearHypothesis(LinearMixed_CFU, c('DAY:ARMTMC207 + PA-824 + PYRAZINAMIDE=DAY:ARMRifafour e-275'))
```

```{r}
linearHypothesis(LinearMixed_PC1, c('DAY:ARMTMC207 + PA-824 + PYRAZINAMIDE=DAY:ARMRifafour e-275'))
```
```{r}

```


#### Coefficients of variation
```{r }
model_list <- c(LinearMixed_CFU,LinearMixed_TTP, LinearMixed_PC1)

calculate_cv <- function(model) {
  re_varcorr <- VarCorr(model)
  random_intercept_var <- as.numeric(re_varcorr$USUBJID[1, 1]) 
  fixed_effect_mean <- fixef(model)[1]
  model_cv <- sqrt(random_intercept_var) / fixed_effect_mean * 100
  return(model_cv)
}


cv_results <- sapply(model_list, calculate_cv)
cv_table <- data.frame(cv_results)
rownames(cv_table) <- c('LinearMixedCFU','LinearMixed_TTP', 'LinearMixed_PC1')
print(cv_table)
```

##### Lollipop/cleveland plots for CV values
```{r}
calculate_cv_FE <- function(model) {
  coef_tab <- coef(summary(model))
  beta<-coef_tab[,"Estimate"]
  se<-coef_tab[,"Std. Error"]
  model_cv <- (se / abs(beta)) * 100
  return(model_cv)
}
cv_results <- sapply(model_list, calculate_cv_FE)
cv_table <- data.frame(cv_results)
colnames(cv_table) <- c('logCFU', 'logTTP', 'PC1')
print(cv_table)
cv_table$arm<-rownames(cv_table)
cv_table<-cv_table[-1,]
```

```{r}
cv_df<-cv_table |>
  pivot_longer(c(logCFU,logTTP,PC1),
    names_to="Assay",
    values_to="CV"
  )
cv_df
```


```{r}
ggplot(cv_df, aes(x=CV, y=arm, color=Assay)) + 
  geom_point(size=3) + 
  geom_segment(aes(y=arm, 
                   yend=arm, 
                   x=0, 
                   xend=max(CV)), linetype="dashed", 
               size=0.1) + 
  labs(title="Cleveland dot chart", 
       subtitle="CV of per-arm TE") + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6)) +
  theme_minimal()
```


### Bilinear mixed models
```{r}

```

