---
title: "pPPCA_model_validation_TBPACTS"
output: html_document
---
## Get pkgs
```{r}
pkgs<-c("tidyverse","lme4","ggplot2","emmeans","RColorBrewer","patchwork","gt",
          "grid","broom.mixed","pbkrtest","jtools","car", "Gmisc", "htmlTable",
        "gtsummary", "lmerTest", "purrr")
lapply(pkgs, require, character.only=T)
```

## Read in data
```{r}
## write out of dose-response models in R
df<-read.csv("/Users/mairi.mcclean/Desktop/btz/manuscript_posters/CPT_PSP_submission/submitted/revisions/extra_analysis/processed_data/260526mm_pPCA_TB1014.csv")
#df<-read.csv("/Users/mairi.mcclean/Desktop/btz/manuscript_posters/CPT_PSP_submission/submitted/revisions/extra_analysis/processed_data/260605mm_pPCA_TB1014_ULQTTP.csv")

#df<-read.csv("/Users/mairi.mcclean/Desktop/btz/manuscript_posters/CPT_PSP_submission/submitted/revisions/extra_analysis/processed_data/260605mm_pPCA_TB1014_ALL_REMOVED.csv")

df<-read.csv("/Users/mairi.mcclean/Desktop/btz/manuscript_posters/CPT_PSP_submission/submitted/revisions/extra_analysis/processed_data/260608mm_pPCA_TB1014_ALL_REMOVED.csv")
df<-df|>
  mutate(across(c(USUBJID,ARM,SEX), as.factor))|>
  rename(PC1 = pc1)|>
  select(-c(logCFU,logTTP))
```

```{r}
df_raw<-read.csv("/Users/mairi.mcclean/Desktop/btz/manuscript_posters/CPT_PSP_submission/submitted/revisions/extra_analysis/processed_data/2600605mm_TB1014_raw.csv")
df_raw<-df_raw|>
  select(c("USUBJID","logCFU","logTTP","DAY","ARM"))
```

##### Drop rows U/BLQ
```{r}
df_cfu<-df_raw|>
  select(c("USUBJID","logCFU","DAY", "ARM"))|>
  filter(!logCFU <= 0) ## Will not keep NAs here - this is correct; will be less than df used for PCA

df_ttp<-df_raw|>
  select(c("USUBJID","logTTP","DAY", "ARM"))|>
  filter(!logTTP >= 2.77)  ## Will not keep NAs here - this is correct; will be less than df used for PCA
```



```{r}
dfs<-list(df,df_cfu,df_ttp)
df_all<-purrr::reduce(.x=dfs,merge, by=c("USUBJID","DAY"),all=T)
```


### Quick vis
```{r}
bm_all<-df_all |>
  pivot_longer(c(logCFU,logTTP,PC1),
    names_to="assay",
    values_to="value"
  ) |>
  select(-c(SEX,AGE,pc2))
```

```{r}
plt<-ggplot(data = bm_all, aes(x = DAY, y = value, color = USUBJID)) +
  geom_line() +
  stat_summary(
    fun = median,
    geom = "line", na.rm = TRUE,
    aes(group = assay, color=assay),
   size = 1.5
  ) + 
#  coord_cartesian()+
  scale_color_manual(
    values = c(
      setNames(rep("grey70", length(unique(bm_all$USUBJID))), unique(bm_all$USUBJID)),
      setNames(brewer.pal(3, "Set2"), unique(bm_all$assay))
    )
  ) +
  facet_wrap(
    ~ assay,shrink = FALSE,
    scales = "free_y",# independent = "y",
    ncol = 3,
             labeller = labeller(
              assay = c("logCFU" = "log10 CFU/ml",
                        "logTTP" = "log10 MGIT TTP (h)",
                        "pc1_all"="PC1")
            )) +
  ggtitle("Per-patient trajectories by measurement") +
  theme_minimal() +
  theme(panel.grid.major = element_blank(),
        legend.position = "none",
        axis.title.x = element_text(size=12),
        axis.title.y = element_text(size=12),
        title = element_text(size = 15),
        legend.text = element_text(size = 15),
        strip.text=element_text(size=12)) +
  labs(y = "Assay value", x = "Days on treament")

plt
```
> MGIT values reaching log10 == 4? this is plotting artifact, no values like this
> exist in df

```{r}
ggsave("260608mm_tb1014_spaghetti_NOULLQ.tiff", plt, device='tiff', dpi=600,
       bg="white",width = 178,height=100,units='mm')
```


### Get summary tables over time
```{r}
### read in the 'raw' data (with the U/LLO values)
#['logCFU'] <= -3.0
#['logTTP'] >= 3.014

X<-read.csv('/Users/mairi.mcclean/Desktop/lva_phII/code_output/tb1014/data_man/250710mm_mb_dm.csv')
X<-X|>
  select(c('USUBJID', 'DAY','ARM', 'logCFU', 'logTTP','AGE', 'SEX'))
```

```{r}
day0<-X|>
  filter(DAY %in% c(-2.0,-1.0))|>
  group_by(USUBJID)|>
  summarise(logCFU=mean(logCFU,na.rm=T),
            logTTP=mean(logTTP,na.rm=T),
            across(c(ARM,SEX,AGE), \(.x) first(.x)))|>
  mutate(DAY = 0)

rest<-X|>
  filter(!DAY %in% c(-2.0,-1.0))

X_<-rbind(day0,rest)

treatment_window<-c(0.0,1.0,2.0,3.0,4.0,5.0,6.0,7.0,8.0,9.0,10.0,11.0,12.0,13.0,14.0)
X_<-X_|>
  filter(DAY %in% treatment_window)
```

```{r}
### get count of values U/LLQ
length(which(X_$logCFU <= log(1))) # 1 CFU/ml <- BURGER
length(which(X_$logTTP >= log(25))) # 25days <- DUFAULT
```


```{r}
X_%>% Gmisc::getDescriptionStatsBy("logCFU", "logTTP", 
                               by = DAY, 
                               add_total_col = F, 
                               hrzl_prop = T, total_col_show_perc = T,
                               show_all_values = T,
                               continuous_fn = describeMedian,
                               header_count = T) %>%
  htmlTable::htmlTable(caption = "Stage 2: Summary statistics by assay over time",
                       align = "center", width = "100%"
                       )%>%
  htmlTable::addHtmlTableStyle(css.cell = "padding-left: .15em; padding-right: .15em;",
                    css.header = "font-weight: normal")

```

```{r}
X_ |>
  tbl_summary(
    by = DAY,
    include = c(logCFU, logTTP),
    digits = all_continuous() ~ 2,
    missing_text = "Missing, n"
  )|>
  modify_header(label="**Assay**")|>
  modify_caption("**Stage 2: Summary statistics by assay over time, per arm**")
```




## Models
### Linear mixed models
#### TTP
```{r}
### Random slope + INTERCEPT
LinearMixed_TTP <- lmer(logTTP ~ DAY:(1+ARM) + (DAY|USUBJID),
                      data = df_ttp) 
coef_model_TTP <- coef(LinearMixed_TTP)$USUBJID

# Now tidy using broom
stage2_slope_ttp <- tidy(LinearMixed_TTP) %>%
  filter(effect == "fixed") %>%
  mutate(ci95 = str_c(
    format(round(estimate,digits=4),nsmall=2),
    " (",
    format(round(estimate - 1.96*std.error,digits=4),nsmall=2),
    ", ",
    format(round(estimate + 1.96*std.error,digits=4),nsmall=2),
    ")"
    )
    
  ) %>%
  dplyr::select(term,ci95)

stage2_slope_ttp$se_fixed_effects <- summary(LinearMixed_TTP)$coefficients[, "Std. Error"]
confint(LinearMixed_TTP, method = "Wald")
# For comparison 
comp_ttp_coefs<- stage2_slope_ttp
```

```{r}
summary(LinearMixed_TTP)
```

```{r}
gr_ttp <- ref_grid(LinearMixed_TTP, cov.keep= c('ARM', 'DAY'))
emm_ttp <- emmeans(gr_ttp, spec= c('ARM', 'DAY'), level= 0.95)
emm_ttp_df <- as.data.frame(emm_ttp)
```

```{r}
ggplot(df, aes(x=DAY, y=logTTP))+
        geom_line(data= data.frame(emm_ttp), aes(y= emmean), color ='blue', size = 1.5) +
        scale_x_continuous(limits = c(0, 15), breaks = c(0,2,4,6,8,10,12,14))+
        facet_wrap(~ARM,nrow=2)+
        labs(y = "log10(TTP) (hrs)",
             x = "Day(s) of treatment")+
        theme(axis.title = element_text(size = 16))

```

#### CFU
```{r}
LinearMixed_CFU <- lmer(logCFU ~ DAY:(1+ARM) + (DAY|USUBJID),
                      data = df_cfu) 
coef_model_CFU <- coef(LinearMixed_CFU)$USUBJID

# Now tidy using broom
stage2_slope_cfu <- tidy(LinearMixed_CFU) %>%
  filter(effect == "fixed") %>%
  mutate(ci95 = str_c(
    format(round(estimate,digits=4),nsmall=2),
    " (",
    format(round(estimate - 1.96*std.error,digits=4),nsmall=2),
    ", ",
    format(round(estimate + 1.96*std.error,digits=4),nsmall=2),
    ")"
    )
    
  ) %>%
  dplyr::select(term,ci95)

stage2_slope_cfu$se_fixed_effects <- summary(LinearMixed_CFU)$coefficients[, "Std. Error"]
confint(LinearMixed_CFU, method = "Wald")
# For comparison 
comp_cfu_coefs<- stage2_slope_cfu
```

```{r}
summary(LinearMixed_CFU)
```

```{r}
gr_cfu <- ref_grid(LinearMixed_CFU, cov.keep= c('ARM', 'DAY'))
emm_cfu <- emmeans(gr_cfu, spec= c('ARM', 'DAY'), level= 0.95)
emm_cfu_df <- as.data.frame(emm_cfu)
```

```{r}
ggplot(df, aes(x=DAY, y=logCFU))+
        geom_line(data= data.frame(emm_cfu), aes(y= emmean), color ='blue', size = 1.5) +
        scale_x_continuous(limits = c(0, 15), breaks = c(0,2,4,6,8,10,12,14))+
        facet_wrap(~ARM,nrow=2)+
        labs(y = "log10(CFU)",
             x = "Day(s) of treatment")+
        theme(axis.title = element_text(size = 16))

```

#### PC1
```{r}
LinearMixed_PC1 <- lmer(PC1 ~ DAY:(1+ARM) + (DAY|USUBJID),
                      data = df) 
coef_model_pc1 <- coef(LinearMixed_PC1)$USUBJID

# Now tidy using broom
stage2_slope_pc1 <- tidy(LinearMixed_PC1) %>%
  filter(effect == "fixed") %>%
  mutate(ci95 = str_c(
    format(round(estimate,digits=4),nsmall=2),
    " (",
    format(round(estimate - 1.96*std.error,digits=4),nsmall=2),
    ", ",
    format(round(estimate + 1.96*std.error,digits=4),nsmall=2),
    ")"
    )
    
  ) %>%
  dplyr::select(term,ci95)

stage2_slope_pc1$se_fixed_effects <- summary(LinearMixed_PC1)$coefficients[, "Std. Error"]
confint(LinearMixed_PC1, method = "Wald")
# For comparison 
comp_pc1_coefs<- stage2_slope_pc1
```
> model converged fine

```{r}
summary(LinearMixed_PC1)
```

```{r}
gr_pc <- ref_grid(LinearMixed_PC1, cov.keep= c('ARM', 'DAY'))
emm_pc <- emmeans(gr_pc, spec= c('ARM', 'DAY'), level= 0.95)
emm_pc_df <- as.data.frame(emm_pc)
```

```{r}
ggplot(df, aes(x=DAY, y=pc1))+
        geom_line(data= data.frame(emm_ttp), aes(y= emmean), color ='blue', size = 1.5) +
        scale_x_continuous(limits = c(0, 15), breaks = c(0,2,4,6,8,10,12,14))+
        facet_wrap(~ARM,nrow=2)+
        labs(y = "log10(TTP) (hrs)",
             x = "Day(s) of treatment")+
        theme(axis.title = element_text(size = 16))

```

```{r}
summary(LinearMixed_CFU)
colnames(coef_model_pc1)
```

```{r}
colnames(coef_model_pc1)
```

#### Forest plot
```{r}
plt <- plot_coefs(LinearMixed_PC1,LinearMixed_TTP,LinearMixed_CFU,
           model.names = c('PC1 (n=105)',
                           'MGIT TTP (n=105)',
                           'CFU (n=105)'),
          omit.coefs = "(Intercept)",
           coefs=c('CLOFAZIMINE'='DAY:ARMCLOFAZIMINE',
                   'PYRAZINAMIDE'='DAY:ARMPYRAZINAMIDE',
                   'Rifafour e-275'='DAY:ARMRifafour e-275',
                   'TMC207 + PA-824 + CLOFAZIMINE'='DAY:ARMTMC207 + PA-824 + CLOFAZIMINE',
                   'TMC207 + PA-824 + PYRAZINAMIDE'='DAY:ARMTMC207 + PA-824 + PYRAZINAMIDE',
                   'TMC207 + PA-824 + PYRAZINAMIDE + CLOFAZIMINE'='DAY:ARMTMC207 + PA-824 + PYRAZINAMIDE + CLOFAZIMINE',
                   'TMC207 + PYRAZINAMIDE + CLOFAZIMINE'='DAY:ARMTMC207 + PYRAZINAMIDE + CLOFAZIMINE'
                   ),
           scale=TRUE, size=8,legend.title="")

plt <- plt +
  scale_color_manual(values = c(
    "CFU (n=105)" = "#F8766D",
    "MGIT TTP (n=105)" = "#7CAE00",
    "PC1 (n=105)" = "#00BFC4"
  )) +
  facet_wrap(~ model, ncol=2, scales = "free_x") +
  theme_minimal() +
#+   scale_x_continuous(breaks = c(0, seq(-1, 1, by = 0.1))) + 
  theme(axis.text.x = element_text(size = 9),
        axis.title.x=element_text(size=12),
        strip.text=element_text(size=10),
        axis.text.y = element_text(size = 10),
#        axis.title.y=element_blank(),
        legend.text=element_text(),
        legend.title = element_text(),
        legend.position='none'
        )+
  guides(shape=guide_legend(reverse = TRUE))+
  labs(x='Model coefficient',
       y='Treatment arm')
#CFU-'#F8766D'
#TTP-'#7CAE00'
#PC1-'#00BFC4'
plt

ggsave("260528mm_linMixMod_CFU_TTP_PC1.png", plt, device='png', 
       dpi = 600, bg='white') 
```

#### Standarised response
##### TTP
```{r}
### Random slope + INTERCEPT
df_ttp$logTTP_sc<- scale(df_ttp$logTTP, center=T,scale=T)

LinearMixed_TTPsc <- lmer(logTTP_sc ~ DAY:(1+ARM) + (DAY|USUBJID),
                      data = df_ttp) 
coef_model_TTP <- coef(LinearMixed_TTPsc)$USUBJID

# Now tidy using broom
stage2_slope_ttp <- tidy(LinearMixed_TTPsc) %>%
  filter(effect == "fixed") %>%
  mutate(ci95 = str_c(
    format(round(estimate,digits=4),nsmall=2),
    " (",
    format(round(estimate - 1.96*std.error,digits=4),nsmall=2),
    ", ",
    format(round(estimate + 1.96*std.error,digits=4),nsmall=2),
    ")"
    )
    
  ) %>%
  dplyr::select(term,ci95)

stage2_slope_ttp$se_fixed_effects <- summary(LinearMixed_TTPsc)$coefficients[, "Std. Error"]
confint(LinearMixed_TTPsc, method = "Wald")
# For comparison 
comp_ttp_coefs<- stage2_slope_ttp
```

##### CFU
```{r}
### Random slope + INTERCEPT
df_cfu$logCFU_sc<- scale(df_cfu$logCFU, center=T,scale=T)

LinearMixed_CFUsc <- lmer(logCFU_sc ~ DAY:(1+ARM) + (DAY|USUBJID),
                      data = df_cfu) 
coef_model_cfu <- coef(LinearMixed_CFUsc)$USUBJID

# Now tidy using broom
stage2_slope_CFU <- tidy(LinearMixed_CFUsc) %>%
  filter(effect == "fixed") %>%
  mutate(ci95 = str_c(
    format(round(estimate,digits=4),nsmall=2),
    " (",
    format(round(estimate - 1.96*std.error,digits=4),nsmall=2),
    ", ",
    format(round(estimate + 1.96*std.error,digits=4),nsmall=2),
    ")"
    )
    
  ) %>%
  dplyr::select(term,ci95)

stage2_slope_CFU$se_fixed_effects <- summary(LinearMixed_CFUsc)$coefficients[, "Std. Error"]
confint(LinearMixed_CFUsc, method = "Wald")
# For comparison 
comp_cfu_coefs<- stage2_slope_CFU
```

##### PC1
```{r}
### Random slope + INTERCEPT
df_all$PC1_sc<- scale(df_all$PC1, center=T,scale=T)

LinearMixed_PCsc <- lmer(PC1_sc ~ DAY:(1+ARM) + (DAY|USUBJID),
                      data = df_all) 
coef_model_PC <- coef(LinearMixed_PCsc)$USUBJID

# Now tidy using broom
stage2_slope_PC <- tidy(LinearMixed_PCsc) %>%
  filter(effect == "fixed") %>%
  mutate(ci95 = str_c(
    format(round(estimate,digits=4),nsmall=2),
    " (",
    format(round(estimate - 1.96*std.error,digits=4),nsmall=2),
    ", ",
    format(round(estimate + 1.96*std.error,digits=4),nsmall=2),
    ")"
    )
    
  ) %>%
  dplyr::select(term,ci95)

stage2_slope_PC$se_fixed_effects <- summary(LinearMixed_PCsc)$coefficients[, "Std. Error"]
confint(LinearMixed_PCsc, method = "Wald")
# For comparison 
comp_PC_coefs<- stage2_slope_PC
```
```{r}
summary(LinearMixed_PCsc)
```

> model converged OK

###### Forest plot
```{r}
plt <- plot_coefs(LinearMixed_TTPsc,LinearMixed_PCsc,LinearMixed_CFUsc,
           model.names = c('MGIT TTP (n=105)','PC1 (n=105)',
                           
                           'CFU (n=105)'),
          omit.coefs = "(Intercept)",
           coefs=c('CLOFAZIMINE'='DAY:ARMCLOFAZIMINE',
                   'PYRAZINAMIDE'='DAY:ARMPYRAZINAMIDE',
                   'Rifafour e-275'='DAY:ARMRifafour e-275',
                   'TMC207 + PA-824 + CLOFAZIMINE'='DAY:ARMTMC207 + PA-824 + CLOFAZIMINE',
                   'TMC207 + PA-824 + PYRAZINAMIDE'='DAY:ARMTMC207 + PA-824 + PYRAZINAMIDE',
                   'TMC207 + PA-824 + PYRAZINAMIDE + CLOFAZIMINE'='DAY:ARMTMC207 + PA-824 + PYRAZINAMIDE + CLOFAZIMINE',
                   'TMC207 + PYRAZINAMIDE + CLOFAZIMINE'='DAY:ARMTMC207 + PYRAZINAMIDE + CLOFAZIMINE'
                   ),
           scale=TRUE, size=8,legend.title="")

plt <- plt +
  scale_color_manual(values = c(
    "CFU (n=105)" = "#F8766D",
    "MGIT TTP (n=105)" = "#7CAE00",
    "PC1 (n=105)" = "#00BFC4"
  )) +
  facet_wrap(~ model, ncol=3, scales = "free_x") +
  theme_minimal() +
#+   scale_x_continuous(breaks = c(0, seq(-1, 1, by = 0.1))) + 
  theme(axis.text.x = element_text(size =12),
        axis.title.x=element_text(size=15, face="bold"),
        strip.text=element_text(size=16, face="bold"),
        axis.text.y = element_text(size = 15, face="bold"),
        axis.title.y = element_text(size = 15, face="bold"),
#        axis.title.y=element_blank(),
        legend.text=element_text(),
        legend.title = element_text(),
        legend.position='none'
        )+
  guides(shape=guide_legend(reverse = TRUE))+
  labs(x='Model coefficient',
       y='Treatment arm')

#CFU-'#F8766D'
#TTP-'#7CAE00'
#PC1-'#00BFC4'
plt
```


```{r}
ggsave("260608mm_linMixMod_CFU_TTP_PC1_standardised_NO_ULQTTP.tiff", plt, device='tiff', 
       width = 18, dpi = 600, units = "in",
       bg='white')#height = 5,
```


####### Table of standardised SEs
```{r}
std_ses<-data.frame(
  tx_arm = comp_PC_coefs$term,
  CFU_std_se = c(comp_cfu_coefs$se_fixed_effects),
  TTP_std_se = c(comp_ttp_coefs$se_fixed_effects),
  PC1_std_se = c(comp_PC_coefs$se_fixed_effects)
)
std_ses
```


```{r}
#MDE <- 2.8 xSE(B_stand) <<- For individual model estimates

```

####### Global precision
```{r}
anova(LinearMixed_TTPsc, type = 3)
anova(LinearMixed_PCsc, type = 3)

cat("TTP Residual Noise:", sigma(LinearMixed_TTPsc)^2, "\n")
cat("PC1 Residual Noise:", sigma(LinearMixed_PCsc)^2, "\n")
```
######## Comparing CI width intervals
```{r}
#'Rifafour e-275'='DAY:ARMRifafour e-275',
#'TMC207 + PA-824 + CLOFAZIMINE'='DAY:ARMTMC207 + PA-824 + CLOFAZIMINE',
#'TMC207 + PA-824 + PYRAZINAMIDE'='DAY:ARMTMC207 + PA-824 + PYRAZINAMIDE',
#'TMC207 + PA-824 + PYRAZINAMIDE + CLOFAZIMINE'='DAY:ARMTMC207 + PA-824 + PYRAZINAMIDE + CLOFAZIMINE',
#'TMC207 + PYRAZINAMIDE + CLOFAZIMINE'='DAY:ARMTMC207 + PYRAZINAMIDE + CLOFAZIMINE'

emm_ttp <- emmeans(LinearMixed_TTPsc, ~ ARM | DAY)
emm_pc1 <- emmeans(LinearMixed_PCsc, ~ ARM | DAY)

combo_contrasts <- list(
  "TMC207 + PA-824 + CLOFAZIMINE vs Rifafour e-275" = c(-1, 1, 0, 0, 0, 0, 0),
  "TMC207 + PA-824 + PYRAZINAMIDE vs Rifafour e-275" = c(-1, 0, 1, 0, 0, 0, 0),
  "TMC207 + PA-824 + PYRAZINAMIDE + CLOFAZIMINE vs Rifafour e-275" = c(-1, 0, 0, 1, 0, 0, 0),
  "TMC207 + PYRAZINAMIDE + CLOFAZIMINE vs Rifafour e-275" = c(-1, 0, 0, 0, 1, 0, 0)
)

# 3. Get Dunnett-adjusted 90% Confidence Intervals
ci_ttp <- confint(contrast(emm_ttp, method = combo_contrasts), level = 0.90, adjust = "mvt")
ci_pc1 <- confint(contrast(emm_pc1, method = combo_contrasts), level = 0.90, adjust ßßß= "mvt")

# 4. Calculate and compare widths
ci_ttp$width <- ci_ttp$upper.CL - ci_ttp$lower.CL
ci_pc1$width <- ci_pc1$upper.CL - ci_pc1$lower.CL

# 5. Print your final evidence table
data.frame(Contrast = ci_ttp$contrast, TTP_Width = ci_ttp$width, PC1_Width = ci_pc1$width)

```



#### ANOVA of model estimates
##### TTP
```{r}
anova(LinearMixed_TTP, type=3)
```

##### CFU
```{r}
anova(LinearMixed_CFU, type=3)
```

##### PC1
```{r}
anova(LinearMixed_PC1, type=3)
```



#### ANOVA of patient estimates
```{r}
## get patient level slopes per measurement for ANOVA
write.csv(coef_model_pc1, "PC_model_est_per_px.csv")
write.csv(coef_model_CFU, "CFU_model_est_per_px.csv")
```

##### TTP
```{r}
write.csv(coef_model_TTP, "TTP_model_est_per_px.csv")
idv_estimates <- read_csv("TTP_model_est_per_px.csv", col_names = TRUE)
names(idv_estimates)[1] <- "Participant.ID"
idv_estimates$USUBJID <- as.factor(idv_estimates$Participant.ID)

```
                                
```{r}
tx_alloc<-df|>
  select(c(USUBJID,ARM))
patient_TTP_estimates <- left_join(idv_estimates,tx_alloc, by = 'USUBJID') %>%
  mutate(ch1 = ifelse(ARM == "CLOFAZIMINE", 1, 0),
         ch2 = ifelse(ARM == "PYRAZINAMIDE", 1, 0),
         ch3 = ifelse(ARM == "Rifafour e-275", 1, 0),
         ch4 = ifelse(ARM == "TMC207 + PA-824 + CLOFAZIMINE", 1, 0),
         ch5 = ifelse(ARM == "TMC207 + PA-824 + PYRAZINAMIDE", 1, 0),
         ch6 = ifelse(ARM == "TMC207 + PA-824 + PYRAZINAMIDE + CLOFAZIMINE", 1, 0),
         ch7 = ifelse(ARM == "TMC207 + PYRAZINAMIDE + CLOFAZIMINE", 1, 0),
         ) %>%
  mutate(idv_slp = DAY + 
           `DAY:ARMCLOFAZIMINE`*ch1 + 
           `DAY:ARMPYRAZINAMIDE`*ch2+
           `DAY:ARMRifafour e-275`*ch3+
           `DAY:ARMTMC207 + PA-824 + CLOFAZIMINE`*ch4+
           `DAY:ARMTMC207 + PA-824 + PYRAZINAMIDE`*ch5+
           `DAY:ARMTMC207 + PA-824 + PYRAZINAMIDE + CLOFAZIMINE`*ch6+
           `DAY:ARMTMC207 + PYRAZINAMIDE + CLOFAZIMINE`*ch7)

write.csv(patient_TTP_estimates,
          "/Users/mairi.mcclean/Desktop/btz/manuscript_posters/CPT_PSP_submission/submitted/revisions/extra_analysis/processed_data/TTP_slopes_per_px.csv")
```


##### CFU
```{r}
write.csv(coef_model_CFU, "CFU_model_est_per_px.csv")
idv_estimates <- read_csv("CFU_model_est_per_px.csv", col_names = TRUE)
names(idv_estimates)[1] <- "Participant.ID"
idv_estimates$USUBJID <- as.factor(idv_estimates$Participant.ID)
```
                                
```{r}
tx_alloc<-df|>
  select(c(USUBJID,ARM))
patient_CFU_estimates <- left_join(idv_estimates,tx_alloc, by = 'USUBJID') %>%
  mutate(ch1 = ifelse(ARM == "CLOFAZIMINE", 1, 0),
         ch2 = ifelse(ARM == "PYRAZINAMIDE", 1, 0),
         ch3 = ifelse(ARM == "Rifafour e-275", 1, 0),
         ch4 = ifelse(ARM == "TMC207 + PA-824 + CLOFAZIMINE", 1, 0),
         ch5 = ifelse(ARM == "TMC207 + PA-824 + PYRAZINAMIDE", 1, 0),
         ch6 = ifelse(ARM == "TMC207 + PA-824 + PYRAZINAMIDE + CLOFAZIMINE", 1, 0),
         ch7 = ifelse(ARM == "TMC207 + PYRAZINAMIDE + CLOFAZIMINE", 1, 0),
         ) %>%
  mutate(idv_slp = DAY + 
           `DAY:ARMCLOFAZIMINE`*ch1 + 
           `DAY:ARMPYRAZINAMIDE`*ch2+
           `DAY:ARMRifafour e-275`*ch3+
           `DAY:ARMTMC207 + PA-824 + CLOFAZIMINE`*ch4+
           `DAY:ARMTMC207 + PA-824 + PYRAZINAMIDE`*ch5+
           `DAY:ARMTMC207 + PA-824 + PYRAZINAMIDE + CLOFAZIMINE`*ch6+
           `DAY:ARMTMC207 + PYRAZINAMIDE + CLOFAZIMINE`*ch7)

write.csv(patient_CFU_estimates,
          "/Users/mairi.mcclean/Desktop/btz/manuscript_posters/CPT_PSP_submission/submitted/revisions/extra_analysis/processed_data/CFU_slopes_per_px.csv")
```


##### PC1
```{r}
write.csv(coef_model_pc1, "PC_model_est_per_px.csv")
idv_estimates <- read_csv("PC_model_est_per_px.csv", col_names = TRUE)
names(idv_estimates)[1] <- "Participant.ID"
idv_estimates$USUBJID <- as.factor(idv_estimates$Participant.ID)

```
                                
```{r}
tx_alloc<-df|>
  select(c(USUBJID,ARM))
patient_PC1_estimates <- left_join(idv_estimates,tx_alloc, by = 'USUBJID') %>%
  mutate(ch1 = ifelse(ARM == "CLOFAZIMINE", 1, 0),
         ch2 = ifelse(ARM == "PYRAZINAMIDE", 1, 0),
         ch3 = ifelse(ARM == "Rifafour e-275", 1, 0),
         ch4 = ifelse(ARM == "TMC207 + PA-824 + CLOFAZIMINE", 1, 0),
         ch5 = ifelse(ARM == "TMC207 + PA-824 + PYRAZINAMIDE", 1, 0),
         ch6 = ifelse(ARM == "TMC207 + PA-824 + PYRAZINAMIDE + CLOFAZIMINE", 1, 0),
         ch7 = ifelse(ARM == "TMC207 + PYRAZINAMIDE + CLOFAZIMINE", 1, 0),
         ) %>%
  mutate(idv_slp = DAY + 
           `DAY:ARMCLOFAZIMINE`*ch1 + 
           `DAY:ARMPYRAZINAMIDE`*ch2+
           `DAY:ARMRifafour e-275`*ch3+
           `DAY:ARMTMC207 + PA-824 + CLOFAZIMINE`*ch4+
           `DAY:ARMTMC207 + PA-824 + PYRAZINAMIDE`*ch5+
           `DAY:ARMTMC207 + PA-824 + PYRAZINAMIDE + CLOFAZIMINE`*ch6+
           `DAY:ARMTMC207 + PYRAZINAMIDE + CLOFAZIMINE`*ch7)

write.csv(patient_PC1_estimates,
          "/Users/mairi.mcclean/Desktop/btz/manuscript_posters/CPT_PSP_submission/submitted/revisions/extra_analysis/processed_data/PC1_slopes_per_px.csv")
```


#### Simple stats on model slopes
```{r}
#### Original authors ran ANOVA (on patient slopes? Or model estimates) TBD
res.ttp<-aov(idv_slp ~ ARM, data=patient_TTP_estimates)
summary(res.ttp)
```

```{r}
res.cfu<-aov(idv_slp ~ ARM, data=patient_CFU_estimates)
summary(res.cfu)
```

```{r}
res.pc1<-aov(idv_slp ~ ARM, data=patient_PC1_estimates)
summary(res.pc1)
```
> PC1 better than CFU, not better than TTP, to determine an overall effect

##### H0 tests
> How are the H0 tests for beta==0, and control == experimental arm
>> Get all 10 and then correct for MTC?

```{r}
# 'DAY:ARMCLOFAZIMINE=0'
linearHypothesis(LinearMixed_TTP, c('DAY:ARMTMC207 + PA-824 + PYRAZINAMIDE=DAY:ARMRifafour e-275'))
```

```{r}
linearHypothesis(LinearMixed_CFU, c('DAY:ARMTMC207 + PA-824 + PYRAZINAMIDE=DAY:ARMRifafour e-275'))
```

```{r}
linearHypothesis(LinearMixed_PC1, c('DAY:ARMTMC207 + PA-824 + PYRAZINAMIDE=DAY:ARMRifafour e-275'))
```


#### Coefficients of variation
```{r }
model_list <- c(LinearMixed_CFU,LinearMixed_TTP, LinearMixed_PC1)

calculate_cv <- function(model) {
  re_varcorr <- VarCorr(model)
  random_intercept_var <- as.numeric(re_varcorr$USUBJID[1, 1]) 
  fixed_effect_mean <- fixef(model)[1]
  model_cv <- sqrt(random_intercept_var) / fixed_effect_mean * 100
  return(model_cv)
}


cv_results <- sapply(model_list, calculate_cv)
cv_table <- data.frame(cv_results)
rownames(cv_table) <- c('LinearMixedCFU','LinearMixed_TTP', 'LinearMixed_PC1')
print(cv_table)
```

##### Lollipop/cleveland plots for CV values
```{r}
calculate_cv_FE <- function(model) {
  coef_tab <- coef(summary(model))
  beta<-coef_tab[,"Estimate"]
  se<-coef_tab[,"Std. Error"]
  model_cv <- (se / abs(beta)) * 100
  return(model_cv)
}
cv_results <- sapply(model_list, calculate_cv_FE)
cv_table <- data.frame(cv_results)
colnames(cv_table) <- c('logCFU', 'logTTP', 'PC1')
print(cv_table)
cv_table$arm<-rownames(cv_table)
cv_table<-cv_table[-1,]
```

```{r}
cv_df<-cv_table |>
  pivot_longer(c(logCFU,logTTP,PC1),
    names_to="Assay",
    values_to="CV"
  )
```


```{r}
ggplot(cv_df, aes(x=CV, y=arm, color=Assay)) + 
  geom_point(size=3) + 
  geom_segment(aes(y=arm, 
                   yend=arm, 
                   x=0, 
                   xend=max(CV)), linetype="dashed", 
               size=0.1) + 
  labs(title="Cleveland dot chart", 
       subtitle="CV of per-arm TE") + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6)) +
  theme_minimal()
```


### Bilinear mixed models
```{r}

```

